Pretty darn good control: when are approximate solutions better than approximate models

### 2023-01-25 manuscript/manuscript.Rmd
added results, moved graphs around, added line graphs to the const mortality tuning figure

@@

@@

@@

>>>>

# Methods

@@

The result of the tuning procedure is shown in Fig. \ref{fig:linear_tuning}. 
There, we see that the optimal value $F^*$ lies in an area with high episode reward variance.
We will see in Fig. \ref{fig:rew_len_fig} that a likely candidate for this variance is the presence of episodes with early endings.
Thus, the value $0.8 F^*$ supports a smaller variance, and correspondingly -- as Fig. \ref{fig:rew_len_fig} shows -- this policy leads to a much smaller probability of a near-extinction event.

```{r linear_tuning, fig.cap = "Result of the tuning procedure for the constant mortality strategy. (A) Mean episode reward (averaged over 50 episodes) shown as scatter data, $2\times$ standard deviation around each scatter point shown as a grey shade. Vertical lines correspond to the optimal constant mortality rate $F^* = 0.465$ (red) and $0.8 F^*=0.0372$ (green). (B) Several constant mortality policies shown for illustrative purposes. Slopes $0.02$, $0.03$, $0.055$ and $0.065$ shown in black; the optimal linear policy $F^*X$ shown in red, while $0.8F^*X$ shown in green.", out.width = "6.5in"}
msy_rewards <- msy |>
  group_by(action, rep) |>
  filter(t == max(t)) |> 
  group_by(action) |>
  summarise(mean_reward = mean(reward),
            sd = sd(reward))

rmsy <- msy_rewards |>
  ggplot(aes(action, mean_reward, 
             ymin = mean_reward - 2*sd,
             ymax=mean_reward+2*sd)) +
  geom_point() + geom_ribbon(alpha=0.4) +
  labs(y = "episode reward", x = "constant mortality") +
  theme(aspect.ratio = 1/1) + 
  geom_vline(aes(xintercept=0.0465), color="red") + 
  geom_vline(aes(xintercept=0.0372), color="green") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# linear function blueprint
linear_f <- function(slope, x) slope*x
# generates linear function with given slope
linear_f_gen <- function(slope) (function(x) slope*x)

# dummy data set to keep ggplot happy
p <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))

# p + stat_function(fun = linear_f_gen(0.5)) + xlim(0,2)
lines <- p + stat_function(fun = linear_f_gen(0.01)) + 
  stat_function(fun = linear_f_gen(0.02)) +
  stat_function(fun = linear_f_gen(0.03)) +
  stat_function(fun = linear_f_gen(0.8*best_action), col = "green") +
  stat_function(fun = linear_f_gen(best_action), col = "red") +
  stat_function(fun = linear_f_gen(0.055)) + 
  stat_function(fun = linear_f_gen(0.065)) +
  xlim(0,2) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

(rmsy | lines) + plot_annotation(tag_levels = c('A'))
```

@@

In Fig. \ref{fig:timeseries} we show one example time series for each of the three controls we analyzed: 
the left panel uses an optimalconstant mortality control $F^*X$, the middle panel uses $0.8 F^* X$, while the right panel is controlled by the trained DRL agent.
We see that that optimal constant mortality control leads to a near-extinction event.
The constant mortality policy at $0.8 F^*$ is already conservative enough to support a sustainable path.
There, the population levels are maintained roughly in a stationary state.
Finally, the DRL time series shows an oscillatory behavior, with a slight dephase between the $Y$ and $Z$ curves, while the $X$ oscillations are roughly anticorrelated with $Z$.

@@

```{r timeseries_old, fig.cap="TBD.", out.width = "6.5in"}
policy_labels <- list("msy" = "Opt. Const. Mortality", "tac"= "80% Opt. Const. Mortality", "ppo"= "DRL")
policy_labeller <- function(variable,value){
  return(policy_labels[value])
}
msy_tac_ppo |> 
    group_by(rep, policy) |>
    filter(rep == 0) |>
    ggplot(aes(t, abundance, col = species)) + 
        geom_line()  + 
        scale_color_manual(values=c('palegreen3', 'coral1', 'deepskyblue4')) +
        facet_wrap(~factor(policy, levels = c("msy", "tac", "ppo")), labeller = policy_labeller ) +
        labs(y = "abundance", x = "time") +
        theme(aspect.ratio = 2/3)
```
